---
title: "Survival Analysis HW1"
author: "Yixiao Sun"
date: "2024-09-29"
output: pdf_document
---



### Question 2
# Question 2 (b)
```{r}
library(survival)
library(dplyr)
library(survminer)
data <- read.csv("~/Desktop/P8124/p8124_hw1/Q2data_extracted-1.csv")
surv_object <- Surv(time = data$Value, event = data$Binary)
km_fit <- survfit(surv_object ~ 1, conf.type = "plain")
km_fit2 <- survfit(surv_object ~ 1, conf.type = "log-log")

summary(km_fit)
summary(km_fit2)
```
Both approaches results' lower and upper confidence bounds falls inside of the [0,1] interval.

# Question 2 (c)
```{r}
plot(km_fit, conf.int = TRUE, main = "Kaplan-Meier Survival Curve",xlab = "Time (days)", ylab = "Survival Chance")
```
# Question 2 (d)
```{r}
km_median <- quantile(km_fit,probs = 0.5)
km_25th <- quantile(km_fit,probs = 0.25)
km_75th <- quantile(km_fit,probs = 0.75)

print(km_median)
print(km_25th)
print(km_75th)

plot(km_fit, conf.int = TRUE, main = "Kaplan-Meier Survival Curve",xlab = "Time (days)", ylab = "Survival Chance")
abline(h = 0.25)
abline(h = 0.5)
abline(h = 0.75)

summary(km_fit)
```
The median is 90. The 25th quantile is 22, the 75th quantile is 180.And based on the KM estimator, the median is 0.431, 25th is 0.706, 75th is 0.173.

# Question 2 (e)
```{r}
CH <- -log(km_fit$surv)
print(CH)
CH_KM <- data.frame(Time = km_fit$time, Cumulative_Hazard = CH)
print(CH_KM)
```


# Question 2 (f)
```{r}
data.frame(Time = km_fit$time, Cumulative_Hazard = cumsum(km_fit$n.event/km_fit$n.risk))
```
# Question 2 (g)
```{r}
# (i)
CH2 <- cumsum(km_fit$n.event/km_fit$n.risk)
CH_NA <-  data.frame(Time = km_fit$time, Cumulative_Hazard = cumsum(km_fit$n.event/km_fit$n.risk))
plot(CH_NA$Time, CH_NA$Cumulative_Hazard, type = "smooth")
#(ii)
plot(log(CH_NA$Time), log(CH_NA$Cumulative_Hazard), type = "smooth")
``` 
Based on the two graphs above, the second plot, the log one, which is using the Weibull model, showing a better linearality compared to the exponential model. Which mean that the Weibull model might be a better choice for this dataset.

# Question 2 (h)
```{r}
FH <- exp(-CH_NA$Cumulative_Hazard)
data.frame(Time = CH_NA$Time,FH)
plot(km_fit,xlab = "Time (days)", ylab = "Survival Chance")
lines(CH_NA$Time,FH)
```
The Fleming-Harrington estimator line is showing on the graph.We have graphed Kaplan-Meier in the questions upper, there aren't big differences between the two lines, so the Fleming_Harrington estimator is on its agreement with the Kaplan-Merier estimator.


